---
title: "Motor Trend Data Analysis"
author: "Calvin Seto"
date: "October 21, 2015"
output: pdf_document
geometry: margin=0.75in
---

# Executive Summary
Most drivers know that manual transmissions get better gas mileage than automatic transmissions.  Some say that automatic transmissions do as well or better with regard to miles per gallon.  There are many theories about the effects of different car specifications on gas mileage.

* In general, cars with more horsepower or displacement mean the car is more powerful, but that doesn't mean they are more fuel efficient.

* A rear axle ratio of 3.42 is associated with everday use cars and better gas mileage with higher speeds.
    + A rear axle ratio of 4.10 is for cars that tow heavy loads with more torque at lower speeds.

* A heavier car is harder to move and less efficient.


We'll perform a data analysis to answer these questions:

* “Is an automatic or manual transmission better for MPG”

* "Quantify the MPG difference between automatic and manual transmissions"

# Exploratory data analysis
See the Appendix for tables 1, 2, and 3 for the counts of cars with each specification in the data set.

The data set contains continuous and categorical variables, specifically the dependent variable, miles per gallon (mpg) and the continuous independent variables weight (wt), displacement (disp), gross horsepower (hp), rear axle ratio (drat), and quarter mile time (qsec).  There are categorical variables, transmission (am 0=automatic, 1=manual) and V-shape or straight engine (vs 0=V, 1=straight).  The remaining variables, number of cylinders (cyl), number of forward gears (gear), and number of carburetors (carb), could be interpreted as either continuous or categorical.  There were no missing values in the data set.

# Model Selection
We started with a model using transmission only to predict mpg, and the coefficient had a low p-value (0.000285), but it's R-squared is also low(0.3598).  Let's to try improve it by adding variables.

We used a correlation table to see how strongly each variable is related to miles per gallon.  We added variables with high correlation to mpg to the model and examined the model's statistics.  To reduce the effects of multicollinearity, we also looked at the correlation amongst each variable.  We omitted variables that were related to each other to reduce the variance of the estimated coefficients, if the predictors were correlated.

Using the top three most correlated variables with mpg, (i.e. wt, cyl, and disp) we created three more models combining each with transmission.  These models had significant p-values for the coefficients for wt, cyl, and disp, high R-squared, and F-test values, but all showed that transmission was NOT significant in predicting miles per gallon.

To control the confounding effects of transmission, we'll design models by restricting and matching the other variables.  Additionally, we'll try an automatic model selection method and the nested modeling/ANOVA technique.  We did not use these models because the transmission coefficient was not significant, the variance inflation factors were too high, indicating the variables were correlated, or other variable's coefficients were not significant.

We arrived at the final model by "Reducing the model" or the practice of including all candidate predictors, and then systematically removing the term with the highest p-value one-by-one until only significant predictors are left.

The R-squared is 0.8497, The Adjusted R-squared is 0.8336 and the Predicted R-squared is 0.7946.

The F-statistic is 52.75 and the p-value is 1.21e-11.

The variance inflation factor for weight is 2.482952, quarter mile time is 1.364339, and transmission is 2.541437.

Figure 1 in the Appendix shows plots of our model's residuals vs fitted values and Normal Q-Q standardized residuals.

# Results
The final parsimonious regression model that easily describes the mean relationship between weight, quarter mile time, transmission and miles per gallon for automatics was
$$
mpg =  9.6178  - 3.9165 wt + 1.2259 qsec 
$$

and for manuals was
$$
mpg =  12.5536  - 3.9165 wt + 1.2259 qsec
$$

The intercept of our model is not significant with a p-value of 0.177915, but we are not interested in the cases where the predictors weight or quarter mile time is zero.  There are statistically significant relationships between weight, quarter mile time, transmission and miles per gallon (p-values of 6.95e-06, 0.000216, and 0.046716).

A change of 1,000 pounds in the weight of the car corresponds to a change of -3.9165 in miles per gallon, holding quarter mile time constant.  A change of 1 sec in the quarter mile time of the car corresponds to a change of 1.2259 miles per gallon, holding the weight of the car constant.  A car with a manual transmission is expected to have a change of 2.9358 miles per gallon.


The 95% confidence intervals of the coefficients of our regression model:

 weight               -5.37333423 -2.459673

 quarter mile time    0.63457320  1.817199

 manual transmission  0.04573031  5.825944


With 95% confidence, we estimate a 1,000 pound change in weight results in a -5.3733 to -2.4597 increase in miles per gallon.

With 95% confidence, we estimate a 1 second change in quarter mile time results in a 0.6346 to 1.8172 increase in miles per gallon.

With 95% confidence, we estimate a manual transmission results in a 0.0457 to 5.8260 increase in miles per gallon.

# Conclusions

Our analysis arrived at a model that adjusts the confounding transmission variable and suggests a linear model relating weight, quarter mile time, and transmission to miles per gallon.

On average, a manual transmission is better than an automatic transmission.

For a car with weight of 2.168 and quarter mile time of 16.8, the average miles per gallon, if using an automatic transmission is 21.72168.  For a manual transmission with the same weight and quarter mile time, the average miles per gallon is 24.65752.  It's estimated that this car with a manual transmission will get 2.93584 more miles per gallon than an automatic.

This is based on a limited and outdated sample of cars that don't represent many of the newer vehicles today which use the technology that makes automatic transmission cars equally or more fuel efficient than manual transmissions.

# Appendix

```{r echo=FALSE}
library(knitr)
am1 <- c("automatics","manuals","")
am2 <- c("19","13","")
vs1 <- c("V engine","straight","")
vs2 <- c("18","14","")
specs1 <- as.data.frame(rbind(am1,am2,vs1,vs2))
names(specs1) <- c("Table 1","","")
row.names(specs1) <- c("Transmission"," ","Engine","  ")
kable(specs1)

cyl1 <- c("4 cyl","6 cyl","8 cyl")
cyl2 <- c("11","7","14")
gear1 <- c("3 gear","4 gear","5 gear")
gear2 <- c("15","12","5")
specs2 <- as.data.frame(rbind(cyl1,cyl2,gear1,gear2))
names(specs2) <- c("Table 2","","")
row.names(specs2) <- c("Cylinders"," ","Gears","  ")
kable(specs2)

carb1 <- c("1 carb","2 carb","3 carb","4 carb","6 carb","8 carb")
carb2 <- c("7","10","3","10","1","1")
specs3 <- as.data.frame(rbind(carb1,carb2))
names(specs3) <- c("Table 3","","","","","")
row.names(specs3) <- c("Carburetors"," ")
kable(specs3)
```

```{r echo=FALSE}
PRESS <- function(linear.model) {
  #' calculate the predictive residuals
  pr <- residuals(linear.model)/(1-lm.influence(linear.model)$hat)
  #' calculate the PRESS
  PRESS <- sum(pr^2)
  
  return(PRESS)
}
pred_r_squared <- function(linear.model) {
  #' Use anova() to get the sum of squares for the linear model
  lm.anova <- anova(linear.model)
  #' Calculate the total sum of squares
  tss <- sum(lm.anova$'Sum Sq')
  # Calculate the predictive R^2
  pred.r.squared <- 1-PRESS(linear.model)/(tss)
  
  return(pred.r.squared)
}

```

## Final Model Regression Analysis
```{r echo=FALSE}
library(datasets)
data(mtcars)
lm21 <- lm(mpg ~ wt + qsec + factor(am), data=mtcars)
summary(lm21)
pred_r_squared(lm21)
library(car)
vif(lm21)

```

## Figure 1 - Residuals Plots
```{r echo=FALSE}
par(mfrow=c(1,2))
plot(lm21,which=1)
plot(lm21,which=2)
```

[The mtcarsAnalysis.Rmd file can be found at my github repo here](https://github.com/calvins/RegressionModels_CourseProject/tree/master/writing)
