# Regression Models Course Project
```{r}
row.names(mtcars)
```


## Some functions to compute predicted R squared
```{r}
PRESS <- function(linear.model) {
  #' calculate the predictive residuals
  pr <- residuals(linear.model)/(1-lm.influence(linear.model)$hat)
  #' calculate the PRESS
  PRESS <- sum(pr^2)
  
  return(PRESS)
}
pred_r_squared <- function(linear.model) {
  #' Use anova() to get the sum of squares for the linear model
  lm.anova <- anova(linear.model)
  #' Calculate the total sum of squares
  tss <- sum(lm.anova$'Sum Sq')
  # Calculate the predictive R^2
  pred.r.squared <- 1-PRESS(linear.model)/(tss)
  
  return(pred.r.squared)
}
model_fit_stats <- function(linear.model) {
  r.sqr <- summary(linear.model)$r.squared
  adj.r.sqr <- summary(linear.model)$adj.r.squared
  pre.r.sqr <- pred_r_squared(linear.model)
  PRESS <- PRESS(linear.model)
  return.df <- data.frame(r.squared = r.sqr, adj.r.squared = adj.r.sqr, pred.r.squared = pre.r.sqr, press = PRESS)
  return(return.df)
}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
```

## Data and library load, integrity check, and summary

```{r}
library(datasets)
library(ggplot2)
library(plyr)
data(mtcars)
sum(is.na(mtcars))
head(mtcars)
summary(mtcars)

```

## Explore the data

```{r}
par(mfrow=c(2,2))
qplot(mpg, geom="histogram",data=mtcars,binwidth=0.5)
qplot(wt, geom="histogram",data=mtcars,binwidth=0.5)
qplot(cyl, geom="histogram",data=mtcars,binwidth=0.5)
qplot(disp, geom="histogram",data=mtcars,binwidth=0.5)
par(mfrow=c(2,2))
qplot(hp, geom="histogram",data=mtcars,binwidth=0.5)
qplot(carb, geom="histogram",data=mtcars,binwidth=0.5)
qplot(qsec, geom="histogram",data=mtcars,binwidth=0.5)
qplot(gear, geom="histogram",data=mtcars,binwidth=0.5)
par(mfrow=c(2,2))
qplot(am, geom="histogram",data=mtcars,binwidth=0.5)
qplot(vs, geom="histogram",data=mtcars,binwidth=0.5)
qplot(drat, geom="histogram",data=mtcars,binwidth=0.5)

```

## wt vs mpg
```{r}
qplot(wt,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## disp vs mpg
```{r}
qplot(disp,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## hp vs mpg
```{r}
qplot(hp,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## qsec vs mpg
```{r}
qplot(qsec,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## drat vs mpg
```{r}
qplot(drat,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## tables
```{r}
table(mtcars$am)
table(mtcars$cyl)
table(mtcars$vs)
table(mtcars$gear)
table(mtcars$carb)

```


## Look at mpg for (0) automatic and (1) manual transmissions
```{r}
boxplot(mpg ~ factor(am), data = mtcars, xlab="transmission", ylab="mpg")
mean(subset(mtcars, am == 0)$mpg)
mean(subset(mtcars, am == 1)$mpg)

```

## lm01 - Fit a model with mpg as the response with just the factor variable transmission
```{r}
lm01 <- lm(mpg ~ factor(am), data = mtcars)
summary(lm01)
pred_r_squared(lm01)
par(mfrow=c(2,2))
plot(lm01)
```

## Correlations
```{r}
sort(cor(mtcars)[1,]) # wt,cyl,disp,hp,carb,qsec,gear,am,vs,drat,mpg
cor(mtcars$cyl,mtcars$disp)
cor(mtcars$wt,mtcars$disp)
cor(mtcars$cyl,mtcars$hp)
cor(mtcars$disp,mtcars$hp)
cor(mtcars$wt,mtcars$cyl)
cor(mtcars$wt,mtcars$hp)

```



## lm02 - Fit model with weight and transmission predicting mpg
```{r}
lm02 <- lm(mpg ~ wt + factor(am), data = mtcars)
summary(lm02)
pred_r_squared(lm02)
library(car)
vif(lm02)
par(mfrow=c(2,2))
plot(lm02)

q02 <- qplot(wt, mpg, data=mtcars, colour=am)
q02 <- q02 + geom_abline(intercept=summary(lm02)$coef[1,1], slope=summary(lm02)$coef[2,1], col="red")
q02

```

## lm03 - Fit model with cylinders and transmission predicting mpg
```{r}
lm03 <- lm(mpg ~ cyl + factor(am), data = mtcars)
summary(lm03)
pred_r_squared(lm03)
library(car)
vif(lm03)
par(mfrow=c(2,2))
plot(lm03)

q03 <- qplot(cyl, mpg, data=mtcars, colour=am)
q03 <- q03 + geom_abline(intercept=summary(lm03)$coef[1,1], slope=summary(lm03)$coef[2,1], col="red")
q03

```


## lm04 - Fit model with disp and transmission predicting mpg
```{r}
lm04 <- lm(mpg ~ disp + factor(am), data = mtcars)
summary(lm04)
pred_r_squared(lm04)
library(car)
vif(lm04)
par(mfrow=c(2,2))
plot(lm04)

q04 <- qplot(disp, mpg, data=mtcars, colour=am)
q04 <- q04 + geom_abline(intercept=summary(lm04)$coef[1,1], slope=summary(lm04)$coef[2,1], col="red")
q04
```

## lm05 - Fit model with wt, cyl, disp, hp and transmission predicting mpg
```{r}
lm05 <- lm(mpg ~ wt + cyl + disp + hp + factor(am), data = mtcars)
summary(lm05)
pred_r_squared(lm05)
library(car)
vif(lm05)
par(mfrow=c(2,2))
plot(lm05)

```

## lm06 - Fit model with wt, hp and transmission predicting mpg
```{r}
lm06 <- lm(mpg ~ wt + hp + factor(am), data = mtcars)
summary(lm06)
pred_r_squared(lm06)
library(car)
vif(lm06)
par(mfrow=c(2,2))
plot(lm06)

```

## lm07 - Fit model with wt, cyl and transmission predicting mpg
```{r}
lm07 <- lm(mpg ~ wt + cyl + factor(am), data = mtcars)
summary(lm07)
pred_r_squared(lm07)
library(car)
vif(lm07)
par(mfrow=c(2,2))
plot(lm07)

```

## lm08 - Fit model with wt interacting with transmission predicting mpg
```{r}
lm08 <- lm(mpg ~ wt * factor(am), data = mtcars)
summary(lm08)
pred_r_squared(lm08)
library(car)
vif(lm08)
par(mfrow=c(2,2))
plot(lm08)
```

## lm09 - Fit model with wt interacting with transmission + qsec predicting mpg
```{r}
lm09 <- lm(mpg ~ wt * am + qsec, data=mtcars)
summary(lm09)
pred_r_squared(lm09)
library(car)
vif(lm09)
par(mfrow=c(2,2))
plot(lm09)

```

## lm10 - Fit weight, transmission, wt interacting with transmission
```{r}
lm10 <- lm(mpg ~ wt + factor(am) + wt * factor(am), data=mtcars)
summary(lm10)
pred_r_squared(lm10)
library(car)
vif(lm10)
par(mfrow=c(2,2))
plot(lm10)

```

## lm11 - Fit weight, transmission, cyl
```{r}
lm11 <- lm(mpg ~ wt + factor(am) + factor(cyl), data=mtcars)
summary(lm11)
pred_r_squared(lm11)
library(car)
vif(lm11)
par(mfrow=c(2,2))
plot(lm11)

```

## lm12 - Fit weight and hp
```{r}
lm12 <- lm(mpg ~ wt + hp, data=mtcars)
summary(lm12)
pred_r_squared(lm12)
library(car)
vif(lm12)
par(mfrow=c(2,2))
plot(lm12)

```

## lm13 - Fit weight and carb
```{r}
lm13 <- lm(mpg ~ wt + carb, data=mtcars)
summary(lm13)
pred_r_squared(lm13)
library(car)
vif(lm13)
par(mfrow=c(2,2))
plot(lm13)

```

## lm14 - Fit weight and disp
```{r}
lm14 <- lm(mpg ~ wt + disp, data=mtcars)
summary(lm14)
pred_r_squared(lm14)
library(car)
vif(lm14)
par(mfrow=c(2,2))
plot(lm14)

```

## lm15 - Fit weight and factor(vs)
```{r}
lm15 <- lm(mpg ~ wt + factor(vs), data=mtcars)
summary(lm15)
pred_r_squared(lm15)
library(car)
vif(lm15)
par(mfrow=c(2,2))
plot(lm15)

```

## lm16 - Fit hp and factor(vs)
```{r}
lm16 <- lm(mpg ~ hp + factor(vs), data=mtcars)
summary(lm16)
pred_r_squared(lm16)
library(car)
vif(lm16)
par(mfrow=c(2,2))
plot(lm16)

```

## lm17 - Fit disp and factor(vs)
```{r}
lm17 <- lm(mpg ~ disp + factor(vs), data=mtcars)
summary(lm17)
pred_r_squared(lm17)
library(car)
vif(lm17)
par(mfrow=c(2,2))
plot(lm17)

```

## Results of stepwise selection using MASS stepAIC
```{r}
library(MASS)
fit<- lm(mpg~cyl+disp+hp+drat+wt+qsec+gear+carb,data=mtcars)
step <- stepAIC(fit, direction="both")
step$anova

```

## lm18 - Fit hp, wt, and cyl
```{r}
lm18 <- lm(mpg ~ hp + wt + cyl, data=mtcars)
summary(lm18)
pred_r_squared(lm18)
library(car)
vif(lm18)
par(mfrow=c(2,2))
plot(lm18)

```

## Some multiple factor models and cyl,gear,carb as continuous variables
```{r}
summary(lm(mpg~wt+factor(am)+factor(vs),data=mtcars))
summary(lm(mpg~wt+factor(am)+factor(cyl),data=mtcars))
summary(lm(mpg~wt+factor(am)+factor(gear),data=mtcars))
summary(lm(mpg~wt+factor(am)+factor(carb),data=mtcars))

summary(lm(mpg~wt+cyl+factor(am),data=mtcars))
summary(lm(mpg~wt+hp+factor(am),data=mtcars))
summary(lm(mpg~wt+gear+factor(am),data=mtcars))
summary(lm(mpg~wt+carb+factor(am),data=mtcars))

```

```{r}
summary(lm(mpg~wt,data=mtcars))$fstatistic
summary(lm(mpg~cyl,data=mtcars))$fstatistic
summary(lm(mpg~disp,data=mtcars))$fstatistic
summary(lm(mpg~hp,data=mtcars))$fstatistic
summary(lm(mpg~drat,data=mtcars))$fstatistic
summary(lm(mpg~vs,data=mtcars))$fstatistic
summary(lm(mpg~am,data=mtcars))$fstatistic
summary(lm(mpg~carb,data=mtcars))$fstatistic
summary(lm(mpg~gear,data=mtcars))$fstatistic
summary(lm(mpg~qsec,data=mtcars))$fstatistic

```


the f statistic decreases with the correlation of each single variable

wt, cyl, disp, hp are all correlated

## Nested modeling
```{r}
nm01 <- lm(mpg ~ wt, data=mtcars)
nm03 <- update(nm01, mpg ~ wt+cyl+disp)
nm05 <- update(nm01, mpg ~ wt+cyl+disp+hp+drat)
nm07 <- update(nm01, mpg ~ wt+cyl+disp+hp+drat+factor(vs)+factor(am))
nm09 <- update(nm01, mpg ~ wt+cyl+disp+hp+drat+factor(vs)+factor(am)+carb+gear)
nm10 <- update(nm01, mpg ~ wt+cyl+disp+hp+drat+factor(vs)+factor(am)+carb+gear+qsec)
anova(nm01,nm03,nm05,nm07,nm09,nm10)

nnm01 <- lm(mpg ~ wt, data=mtcars)
nnm03 <- update(nnm01, mpg ~ wt+disp+hp)
nnm05 <- update(nnm01, mpg ~ wt+disp+hp+drat+qsec)
anova(nnm01,nnm03,nnm05)

nnnm01 <- lm(mpg ~ wt, data=mtcars)
nnnm03 <- update(nnm01, mpg ~ wt+cyl+disp)
nnnm05 <- update(nnm01, mpg ~ wt+cyl+factor(am)+hp+drat)
anova(nnnm01,nnnm03,nnnm05)

```

## lm19 - Fit wt, disp and hp
```{r}
lm19 <- lm(mpg ~ wt + disp + hp, data=mtcars)
summary(lm19)
pred_r_squared(lm19)
library(car)
vif(lm19)
par(mfrow=c(2,2))
plot(lm19)

```

## lm20 - Fit wt, disp and hp and transmission
```{r}
lm20 <- lm(mpg ~ wt + disp + hp + factor(am), data=mtcars)
summary(lm20)
pred_r_squared(lm20)
library(car)
vif(lm20)
par(mfrow=c(2,2))
plot(lm20)

```

# diagnostics
```{r}
sort(round(dfbetas(lm20),3))

sort(round(hatvalues(lm20),3))

pairs(mtcars)

```

# Start with full model, remove variables with high p-values until all coefs significant
```{r}
full <- lm(mpg~cyl+disp+hp+drat+wt+qsec+factor(vs)+factor(am)+gear+carb,data=mtcars)
summary(full)
summary(lm(mpg~disp+hp+drat+wt+qsec+factor(vs)+factor(am)+gear+carb,data=mtcars))
summary(lm(mpg~disp+hp+drat+wt+qsec+factor(am)+gear+carb,data=mtcars))
summary(lm(mpg~disp+hp+drat+wt+qsec+factor(am)+gear,data=mtcars))
summary(lm(mpg~disp+hp+drat+wt+qsec+factor(am),data=mtcars))
summary(lm(mpg~disp+hp+wt+qsec+factor(am),data=mtcars))
summary(lm(mpg~hp+wt+qsec+factor(am),data=mtcars))
summary(lm(mpg~wt+qsec+factor(am),data=mtcars))
plot(lm(mpg~wt+qsec+factor(am),data=mtcars))

```


## lm21 - Fit wt, qsec and transmission
```{r}
lm21 <- lm(mpg ~ wt + qsec + factor(am), data=mtcars)
summary(lm21)
pred_r_squared(lm21)
library(car)
vif(lm21)
par(mfrow=c(2,2))
plot(lm21)

```
the intercept is the response when the predictors are 0.  we are not interested in wt and qsec = 0

## lm22 - Fit wt, qsec and transmission; try centering mpg, wt and qsec due to insignificant intercept
```{r}
lm22 <- lm(mpg~I(wt-mean(wt))+I(qsec-mean(qsec))+factor(am), data=mtcars)
summary(lm22)
pred_r_squared(lm22)
library(car)
vif(lm22)
par(mfrow=c(2,2))
plot(lm22)

```

