---
title: "Final mtcars analysis"
author: "Calvin Seto"
date: "October 21, 2015"
output: html_document
---
## Some functions to compute predicted R squared
```{r}
PRESS <- function(linear.model) {
  #' calculate the predictive residuals
  pr <- residuals(linear.model)/(1-lm.influence(linear.model)$hat)
  #' calculate the PRESS
  PRESS <- sum(pr^2)
  
  return(PRESS)
}
pred_r_squared <- function(linear.model) {
  #' Use anova() to get the sum of squares for the linear model
  lm.anova <- anova(linear.model)
  #' Calculate the total sum of squares
  tss <- sum(lm.anova$'Sum Sq')
  # Calculate the predictive R^2
  pred.r.squared <- 1-PRESS(linear.model)/(tss)
  
  return(pred.r.squared)
}
model_fit_stats <- function(linear.model) {
  r.sqr <- summary(linear.model)$r.squared
  adj.r.sqr <- summary(linear.model)$adj.r.squared
  pre.r.sqr <- pred_r_squared(linear.model)
  PRESS <- PRESS(linear.model)
  return.df <- data.frame(r.squared = r.sqr, adj.r.squared = adj.r.sqr, pred.r.squared = pre.r.sqr, press = PRESS)
  return(return.df)
}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
```

## Data and libraries load, integrity check, and summary

```{r}
library(datasets)
library(ggplot2)
library(plyr)
data(mtcars)
sum(is.na(mtcars))
head(mtcars)
summary(mtcars)

```

## Exploratory analysis
```{r}
row.names(mtcars)
```


```{r}
par(mfrow=c(2,2))
qplot(mpg, geom="histogram",data=mtcars,binwidth=0.5)
qplot(wt, geom="histogram",data=mtcars,binwidth=0.5)
qplot(cyl, geom="histogram",data=mtcars,binwidth=0.5)
qplot(disp, geom="histogram",data=mtcars,binwidth=0.5)
par(mfrow=c(2,2))
qplot(hp, geom="histogram",data=mtcars,binwidth=0.5)
qplot(carb, geom="histogram",data=mtcars,binwidth=0.5)
qplot(qsec, geom="histogram",data=mtcars,binwidth=0.5)
qplot(gear, geom="histogram",data=mtcars,binwidth=0.5)
par(mfrow=c(2,2))
qplot(am, geom="histogram",data=mtcars,binwidth=0.5)
qplot(vs, geom="histogram",data=mtcars,binwidth=0.5)
qplot(drat, geom="histogram",data=mtcars,binwidth=0.5)

```

## wt vs mpg
```{r}
qplot(wt,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

### Scatter plots

## disp vs mpg
```{r}
qplot(disp,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## hp vs mpg
```{r}
qplot(hp,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## qsec vs mpg
```{r}
qplot(qsec,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## drat vs mpg
```{r}
qplot(drat,mpg,data=mtcars,geom=c("point","smooth"),method="loess")

```

## tables
```{r}
table(mtcars$am)
table(mtcars$cyl)
table(mtcars$vs)
table(mtcars$gear)
table(mtcars$carb)

```

### Correlations
sort(cor(mtcars)[1,]) # wt,cyl,disp,hp,carb,qsec,gear,am,vs,drat,mpg
cor(mtcars$cyl,mtcars$disp)
cor(mtcars$wt,mtcars$disp)
cor(mtcars$cyl,mtcars$hp)
cor(mtcars$disp,mtcars$hp)
cor(mtcars$wt,mtcars$cyl)
cor(mtcars$wt,mtcars$hp)

## Modeling
## lm21 - Fit wt, qsec and transmission
```{r}
library(datasets)
data(mtcars)
lm21 <- lm(mpg ~ wt + qsec + factor(am), data=mtcars)
summary(lm21)
pred_r_squared(lm21)
library(car)
vif(lm21)
par(mfrow=c(1,2))
plot(lm21,which=1)
plot(lm21,which=2)
# round(dfbetas(lm21),3)
summary(round(dfbetas(lm21),3))
# round(hatvalues(lm21),3)
summary(round(hatvalues(lm21),3))

```


## Get estimates and confidence intervals
```{r}
confint(lm21)
# sumCoef <- summary(lm21)$coefficients
# sumCoef[1,1] + c(-1, 1) * qt(.975, df = lm21$df) * sumCoef[1, 2]
# sumCoef[2,1] + c(-1, 1) * qt(.975, df = lm21$df) * sumCoef[2, 2]
# sumCoef[3,1] + c(-1, 1) * qt(.975, df = lm21$df) * sumCoef[3, 2]
# sumCoef[4,1] + c(-1, 1) * qt(.975, df = lm21$df) * sumCoef[4, 2]

predict(lm21,newdata=data.frame(wt=2.168,qsec=16.8,am=0))
predict(lm21,newdata=data.frame(wt=2.168,qsec=16.8,am=1))

```

## Figure making

